Clustering based on probability distributions with application on residential customers

### 2021-09-29 main.Rmd
adding Characterization of clusters

_Pre-processing steps_

- Handling trend, seasonality, non-stationarity and auto-correlation: Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality. These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method. There is no need to omit holiday or weekend patterns for similar reasons.

- Data transformation: Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.

@@

>>>>

- _Pre-processing steps_

    **Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality. These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method. There is no need to omit holiday or weekend patterns for similar reasons.

    **Data transformation:** Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.

@@

Depending on the distance measure utilized for the study, the cluster characterization technique will differ. Clusters that utilise intra-category distances are characterised using multi-dimensional scaling and parallel coordinate displays. For inter-category distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met. This technique may potentially make advantage of multi-dimensional scaling.

\noindent Multidimensional scaling (MDS) (XXX reference) refers to a family of methods that analyse a matrix of distances or dissimilarities to provide a representation of the data points in a reduced-dimension space. There are many kinds of MDS, but they all solve the same fundamental issue: Given a $n*n$ matrix of dissimilarities and a distance measure, identify a configuration of $n$ points $x_1, x_2, \dots, x_n$ in the reduced dimension space $R^q$ ($q<p$) where the distance between the points is near to the dissimilarity between the points. All techniques must determine the coordinates of the points as well as the space dimension, $q$. Metric and nonmetric MDS are the two main kinds of MDS. Metric MDS methods presume a functional connection between the interpoint distances and the supplied dissimilarities and assume that the data are quantitative. We use metric MDS.

\noindent Parallel coordinate plots (XXX reference) Parallel coordinates have been extensively used to display high-dimensional and multivariate data, allowing for the detection of patterns within the data via visual grouping.
